```{r Environment Setup, include=FALSE}
library(quantmod)
options("getSymbols.warning4.0"=FALSE)

library(FinancialInstrument)
# The package as a whole is abbreviated as "FI" within this document.

# Need to initialize FI for the current we care about.
currency("USD")

library(envvar)
#mailR_conf <- env_or_default("MAILR_CONF", "~/.mailR")
#source(mailR_conf)
```
# Financial Securities Summary
* **By:** David G. Johnston (david.g.johnston@gmail.com)
* **GitHub:** polobo/FinancialSecuritiesSummary

## Purpose

This document contains selected summary information and graphics
for multiple user-specified securities.  Nothing particularly
special and the only "quant" is the calculation of a 120-day moving
average.  This may morph into something more useful than a daily
report but for the moment that is the only goal.  I was getting annoyed
at not having a compact form to view this information on a daily basis.

## Limitations

Name resolution is currently an in-progress work.  If the supplied security
has a ticker symbol that Yahoo! can resolve a length-limited name will be
returned by that service and the FinancialInstrument database addition
should suceed with adding the security as a `stock` type.  This applies to
at least some Vanguard mutual funds.

Within the `updateFIDatabase` function you can add entries to a symbol/name
override routine and these will then be displayed throughout the rest of the
documentation.

Spent way too much time trying to figure out how to suppress the `[1] "Symbol Name"`
output from the Yahoo! meta-data lookup in FI.  Seems to an unfavorable interaction
between `knitr` and `FinancialInstruments`.

"USD" is hard-coded as the currency.

```{r Functions, include=FALSE}
# Print to output the supplied name and the most recent close and ema200
# value from the full xts object passed to the function.
# Called for its side-effects.
printCloseVsEMA <- function(symbolName, closexts, ema200xts) {
  # Pull out the last xts entry and also capture the numeric values
  lastEMA200 <- last(ema200xts)
  lastEMA200Value <- coredata(lastEMA200)
  lastClose  <- last(closexts)
  lastCloseValue <- coredata(lastClose)

  # Proceed to write directly to output; including explicit newlines
  cat(symbolName, "\n")
  cat("EMA200:", round(lastEMA200, digits = 2), "\n")
  cat("CLOSE :", lastClose) #intentional missing newline to append ratio on same line

  # We express the ratio of the current close and ema200 values as a percentage
  cat(" (", round((((lastCloseValue / lastEMA200Value) - 1) * 100), digits = 2), "% )")
}

# A helper function to invisibly update the names found within the
# FI database for specific symbols.
# Called for its side-effects.
updateNameIfExists <- function(symbol, newName) {
  # silently check for the symbol
  if( instrumentExists(symbol) ) {
    instrument_attr(symbol, "name", newName)
  }
}

# Provide an "exists" function for FinancialInstrument since it doesn't
instrumentExists <- function(instrumentName) {
  fi <- getInstrument(instrumentName, silent=TRUE)
  # FI is stupid!!! Instead of returning "NA" if a symbol does not exist
  # instead it returns logical FALSE; but a found symbol returns a list.
  # So we are forced to resort to type checking as a presence indicator.
  # There does not appear to be a built-in "exists" function in FI
  ifelse ( typeof(fi) == "logical", FALSE, TRUE )
}

# The quantmod package does not handle symbol labelling; only quantative data
# So we construct an in-memory database of the symbols that are seen and
# populate the symbol's meta-data within FinancialInstruments package.
# Called for its side-effects.
updateFIDatabase <- function(symbol) {
  stock(symbol, "USD")

  # TODO: this outputs [#] Name to output
  # and while the normal suppression mechanics work interactively
  # something about the interaction with knitr is causing it to appear
  # in the result document no matter what I try surrounding the function
  # call itself.
  update_instruments.yahoo(symbol)

  cat("Yahoo! Resolved", symbol, "to", print(getInstrument(symbol)$name), "\n")

  # TODO: Yahoo! has a length limitation on the "name" attribute
  # so we probably want to setup some kind of override map or
  # choose a different meta-data service.
  # NB: The searching is local and the number of entries is expected to be small
  # so just do the FI-DB check on each new symbol for now.
  updateNameIfExists("VCR", "Vangaurd Consumer Discretionary ETF [VCR]")
}

# Encapsulate the calculations and output desired for each
# individual security.  This consists of both a textual summary
# and one or more charts.
# Called for its side-effects.
displaySymbol <- function(symbol) {
  symboldata <- getSymbols(symbol, env=NULL)

  closexts  <- Cl(symboldata)
  ema200xts <- EMA(closexts, n = 200)

  updateFIDatabase(symbol)
  printCloseVsEMA(getInstrument(symbol)$name, closexts, ema200xts)

  chob <- chartSeries(symboldata, subset="last 36 months", plot=FALSE)
  plot(addTA(ta = ema200xts, col=6, on=1))

  chob <- chartSeries(symboldata, subset="last 14 days", plot=FALSE)
  plot(addTA(ta = ema200xts, col=6, on=1))
}
```

## Portfolio

```{r Print Portfolio}
# Here we actually identify the securities to print
displaySymbol("VCR")
displaySymbol("VFH")
displaySymbol("VIS")
```

## Watch List

```{r Print Watch List}
displaySymbol("VNQ")
displaySymbol("VAW")
displaySymbol("VDE")
displaySymbol("VOX")
```

## Summary

Now that individual securities have been described we can gather up the contents
of FinancialInstruments database and the attached attributes and xts objects to
create an online table and/or a CSV file which can be copied into a custom Excel
model for additional manipulation.  Eventually some of the Excel stuff will
probably migrate to here...

**Not Implemented Yet!**
